---
title: "FARSanalyzr_vignette"
author: "Michal Rackiewicz"
date: "2024-03-03"
output: rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
---

# Introduction

The `FARSanalyzr` package provides functions to read and analyze data from the Fatality Analysis Reporting System (FARS). This vignette provides an overview of the package and demonstrates how to use its main functions.

# Installation

You can install the `FARSanalyzr` package from GitHub using the `devtools` package:

```{r}
devtools::install_github("Karrrramba/Advanced_R_FARS")
```

# Reading FARS data
To read FARS data for a specific year, you can use the fars_read_years function:
```{r}
library(FARSanalyzr)

# Read FARS data for years 2013 and 2014
fars_data <- fars_read_years(c(2013, 2014))
```

# Summarizing FARS data
Once you have read the FARS data, you can summarize it using the fars_summarize_years function:
```{r}
# Summarize total accidents for each month for the years 2018 and 2019
summary_data <- fars_summarize_years(c(2013, 2014))
```

# Plotting 
You can also make a geospatial representation of the recorded events for a given 
state and year.
```{r}
# Map the accidents in the state of California in the year 2019
fars_map_state(6, 2013)
```

